### 2017-09-21 Article analysis.Rmd
Beginnings

---
title: "Article analysis"
author: "Maria C. Codlin"
date: "September 21, 2017"
output: html_document
---
Before I run any parts of the analysis, I load all the packages I use in this file
```{r load libraries}
require(dplyr)
require(ggplot2)
```
First in excel, I deleted two rows of non-table data, i.e. title, and a note, and saved the file (S2_Table) as a CSV in the project directory. Then I read it into R and converted it to a table using dplyr
```{r readfile}
iso<-read.csv("S2_Table.csv")
iso<-tbl_df(iso)
head(iso)
```

To make life easier, I changed the names of two key variables. First I tried 18O and 13C as the variable names, but it through me an error, I think by starting with a number. 
```{r }
names(iso)[12]<-"C13ap"
names(iso)[13]<-"O18ap"
names(iso)
```
This is my first attempt at a box plot! I figured it out eventually
```{r }
g<-ggplot(iso, aes(x=Site, y=O18ap))
g+geom_boxplot(aes(color=Genus))
```

```{r }

```