### 2017-10-07 Article analysis.Rmd

---
title: "Article analysis"
author: "Maria C. Codlin"
date: "September 21, 2017"
output: html_document
---
###Table of contents?

###Introduction to this document, the contents, and the article.

Before I run any parts of the analysis, I load all the packages I use in this file
```{r load libraries}
require(dplyr)
require(stringr)
require(ggplot2)
require(qwraps2)
```
First in excel, I deleted two rows of non-table data, i.e. title, and a note, and saved the file (S2_Table) as a CSV in the project directory. Then I read it into R and converted it to a table using dplyr
```{r readfile}
iso<-read.csv("S2_Table.csv")
iso<-tbl_df(iso_total)
head(iso_total)
```

To make life easier, I changed the names of two key variables. First I tried 18O and 13C as the variable names, but it threw me an error, I think by starting with a number. 
```{r }
names(iso)[12]<-"C13ap"
names(iso)[13]<-"O18ap"
names(iso)
```
This is my first attempt at a box plot! I figured it out eventually
```{r }
g<-ggplot(iso, aes(x=Site, y=O18ap))
g+geom_boxplot(aes(color=Genus))+labs(x="" ,y="δ18O apatite-VSMOW (‰)")
```
Continuation of cleaning. Some samples were removed from analysis due to the likely hood of contamination or diagenesis. " Five apatite samples with C/P ratios <0.10, three apatite samples with IR-SF values < 2.00, and five apatite samples with IR-SF values >4.00 were excluded from subsequent discussions"

```{r }
iso<-filter(iso_total, between (C.P, 0.1, 0.5)&between (IR.SF, 2.0, 4.0))
iso            
```

###Outline of what needs to be done. 
Datasets: 1. Isotopes of modern specimens. 2. Isotopes of ancient specimens.

##Modification of dataset prior to analysis
- calibration of carbon isotope ratios due to atmospheric changes in the past 130 years to make the modern specimens in the dataset comparable to the ancient ones.**It is not clear whether this is incorperated in the supplimental data or not**

##Statistical analyses:
- Parametric tests: Independent samples t-test (groups of two) and ANOVA (and two or more)
-Non parametic: Mann-Whitney U test (for Oztoyahualco and pooled other areas) due to very different sample sizes.
-To see if carbon and oxygen isotopes covary, bivariate Pearson's correlation was used. Sig assumed at a=0.05.

##Mixing model to approximate percentage of C4 and CAM plants in diet:
%C4CAM= ((delta_13_C_apatite - 9.7)-(-26.1))/14.1
Due to the level of uncertainties, these percentages are used to reflect the relative contrabutions of C4 and CAM plants between individuals rather than as absolute data.

```{r summary}
summarise(iso)

```

*Because of errors later on, I found I also needed to trim the space at the end of some of the Genus names.* 
```{r trim white space}
iso_total$Genus<-str_trim(iso_total$Genus)
iso_total$Genus<-as.factor(iso_total$Genus)
levels(iso_total$Genus)
```

```{r sum_table}
my_sum <- 
  list("Archaeological samples" = 
         list("δC13 mean (sd)" = ~ mean_sd(C13ap),
              "δO18 mean (sd)" = ~ mean_sd(O18ap)),
       "Modern samples" = 
         list("δC13 mean (sd)" = ~ mean_sd(modern$modC13),
              "δO18 mean (sd)" = ~ mean_sd(modern$modO18))
              )

summary_table(iso, my_sum[1])
summary_table(iso, my_sum[2])
```

```{r summary}
sum(!is.na(iso$C13ap))#the ! means it will return the number that is not NA
sum(!is.na(iso$O18ap))
arch_sum<-summarize(iso, meanC13=mean(C13ap, na.rm=TRUE), meanO18=mean(O18ap, na.rm=TRUE), sdC13=sd(C13ap, na.rm=TRUE), sdO18=sd(O18ap, na.rm=TRUE))%>%print()

mod_sum<-summarize(modern, meanC13=mean(modC13, na.rm=TRUE), meanO18=mean(modO18, na.rm=TRUE), seC13=std.error(modC13, na.rm=TRUE), sdO18=sd(modO18, na.rm=TRUE))%>%print()

```

These results also differ from what was presented. So I will test to see if they removed outliers prior to t-test. I will use a box plot to identify outliers
```{r}
boxplot(SC, ylim = c(), main = "dC13 ap", xlab = "Sylvilagus")
boxplot(LC, ylim = c(), main = "dC13 ap", xlab = "Lepus")
boxplot(SO, ylim = c(), main = "dO18 ap", xlab = "Sylvilagus")
boxplot(LO, ylim = c(), main = "dO18 ap", xlab = "Lepus")
```