### 2017-10-07 summary stats 2.Rmd

```{r sum_table}
my_sum <- 
  list("Lepus" = 
        round (c("δC13 n" = sum(!is.na(iso$C13ap)),
              "δC13 mean" =  mean(iso$C13ap, na.rm=TRUE),
              "δC13 sd" =  sd(iso$C13ap, na.rm=TRUE),
              "δO18 n" = sum(!is.na(iso$O18ap)),
              "δO18 mean" =  mean(iso$O18ap, na.rm=TRUE),
              "δO18 sd" =  sd(iso$O18ap, na.rm=TRUE)),2),
       "Sylvilagus" = 
         round(c("δC13 n" = sum(!is.na(modern$modC13)),
              "δC13 mean" =  mean(modern$modC13, na.rm=TRUE),
              "δC13 sd" =  sd(modern$modC13, na.rm=TRUE),
              "δO18 n" = sum(!is.na(modern$modO18)),
              "δO18 mean" =  mean(modern$modO18, na.rm=TRUE),
              "δO18 sd" =  sd(modern$modO18, na.rm=TRUE)),2)
       )

my_sum
```

```{r}
sum2<-function(x, y){
for(i in levels(iso$Site)){
  for(i in levels(iso$Genus)){
  x1 <- with(iso$Genus, x[Genus==i])
  y1 <- with(iso$Genus, y[Genus==i])
result<- summarise(mean (x1, na.rm=TRUE))

  }
  return(result)}
}

```
```{r}
sum2(x=iso$C13ap, y=iso$O18ap)
```

round (c("δC13 n" = sum(!is.na(x1)),
              "δC13 mean" =  mean(x1, na.rm=TRUE),
              "δC13 sd" =  sd(x1, na.rm=TRUE),
              "δO18 n" = sum(!is.na(y1)),
              "δO18 mean" =  mean(y1, na.rm=TRUE),
              "δO18 sd" =  sd(y1, na.rm=TRUE)),2)

```{r}
library(dplyr)
library(tidyr)
iso %>%
  group_by(Genus, Site) %>%
  summarize(meanC13=mean(C13ap, na.rm=TRUE)) %>%
  spread(Genus, meanC13)
```
```{r}
carbon_group<-iso %>%
  select(lab_n:Genus, C13ap)%>%
  filter(!(is.na(C13ap)))%>%
  group_by(Site, Genus)
carbon_group
oxygen_group<-iso %>%
  select(lab_n:Genus, O18ap)%>%
  filter(!(is.na(O18ap)))%>%
  group_by(Site, Genus)

carbon_group
oxygen_group
```
```{R}
summ<-carbon_group%>%
  summarize(length(C13ap), mean(C13ap), sd(C13ap))
summ2<-oxygen_group %>%
  summarize(length(O18ap), mean(O18ap), sd(O18ap))

M<-merge(summ,summ2)
M[,3:8]<-round(M[3:8], 2)
M
sum(M$`length(C13ap)`)

```
round(M, 2)
```
t1<-iso$C13ap %>%
  group_by(Site, Genus) %>%
  count()
t2<-iso$O18ap %>%
  group_by(Site, Genus) %>% 
  length()
summ
t1
m1<-cbind(summ, t1)
m2<-cbind(summ2, t2)
m1
merge(m1, m2)
```

```{r}
m<-NULL

iso %>%

 for(i in levels(iso$Genus)){
   tmp <- with(iso, C13ap[Genus=="Lepus"])
        m$δC13_n=sum(!is.na(tmp))

       }

```